Relationships in the extremes and their influence on competition and coexistence

### 2022-07-12 paper/Paper.Rnw
switching paragraphs 1 and 2

%Intro to competition, coexistence, MCT, etc.
Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
the other and drive it extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
in competition'' and ``storage effect'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
%Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
Though MCT's original analytical form presented mathematical challenges for applications to and tests of new systems, recent extensions expand testability and accessibility \citep{Ellner_2016,Ellner_2019}.

@@

%Intro to storage effects and why they might be influenced by ATAs
It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular: the storage effect. We review, conceptually, 
what the storage effect is \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
Storage effect allows competing species to fluctuate %over time while escaping extinction because storage effects provide 
in abundance without any going extinct by providing
a relative benefit to whichever species is currently rare; this is the characteristic that makes the storage effect a so-called \emph{stabilizing} mechanism 
of coexistence \citep{Chesson_2000}.
Storage effect can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
covariation between the fluctuating quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance), that is stronger when the species is common than when it is rare. 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. 
Third, a buffering mechanism %must exist whereby species limit 
which limits the influence of competition %in times 
when the environment is poor. 
Dormant stages such as seed banks are an example buffering mechanism \citep{Chesson_2012}. 
When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
Importantly, the storage effect can be quantified by measuring to what extent EC covariance contributes to differences between 
rare and common species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
Thus ATAs may, in principle, influence storage effect simply because storage effect hinges on EC covariation, and ATAs, when they 
occur between environment and competition variables, can be an aspect of EC covariation.

>>>>

@@

%Intro to competition, coexistence, MCT, etc.
Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
the other and drive it extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] %is a 
has emerged as a
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
MCT breaks down coexistence into the contributions of various mechanisms, %with names such as ``relative nonlinearityin competition'' and ``storage effect'' \citep{Chesson_1994} which will be familiar to many readers. 
highlighting the importance of environmental fluctuations to species coexistence. 
For example, one such mechanism, the storage effect, promotes coexistence when species can reap or "store" the benefits of times when environmental conditions are good to overcome competition and persist through bad conditions.
MCT has been applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
%Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
Though MCT's original analytical form presented mathematical challenges for applications to and tests of new systems, recent developments expand its testability and accessibility and present the opportunity for further extensions \citep{Ellner_2016,Ellner_2019}.

%Intro to storage effects and why they might be influenced by ATAs
In this paper, we show how ATAs can also be important for competition and for the coexistence of
competing species. In particular, we show how ATAs can strongly influence one specific coexistence mechanism:
storage effects.
%It is straightforward to see, intuitively, why ATAs may influence one mechanism of coexistence in particular: the storage effect. 
%We review, conceptually, what the storage effect is \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
Storage effect allows competing species to fluctuate %over time while escaping extinction because storage effects provide 
in abundance without any going extinct by providing
a relative benefit to whichever species is currently rare; this is the characteristic that makes the storage effect a so-called \emph{stabilizing} mechanism 
of coexistence \citep{Chesson_2000}.
Storage effect can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
covariation between the fluctuating quality of the environment (E)
for a species and the strength of competition (C) experienced by that species (called EC covariance), that is stronger when the species is common than when it is rare. 
Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
same environmental conditions to be optimal. 
Third, a buffering mechanism %must exist whereby species limit 
which limits the influence of competition %in times 
when the environment is poor. 
Dormant stages such as seed banks are an example buffering mechanism \citep{Chesson_2012}. 
When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
Importantly, the storage effect can be quantified by measuring to what extent EC covariance contributes to differences between 
rare and common species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
Thus ATAs may, in principle, influence storage effect simply because storage effect hinges on EC covariation, and ATAs, when they 
occur between environment and competition variables, can be an aspect of EC covariation.
